---
title: "Team 5 EDA"
format: html
editor: visual
---

## EDA Team5

```{r}
library(dplyr)
library(readxl)
```

```{r}
Happiness_Data <- read_xls("data/DataFortable2.1WHR2023.xls")
Mental_Health_Data <- read_xlsx("data/Mental health Depression disorder Data.xlsx")
```

```{r}
glimpse(Happiness_Data)
```

```{r}
glimpse(Mental_Health_Data)
```

```{r}
# Show unique values of a column
unique_values_happiness <- unique(Happiness_Data$`Country name`)

# Print the unique values
print(unique_values_happiness)
```

```{r}
# Show unique values of a column
unique_values_mental <- unique(Mental_Health_Data$Entity)

# Print the unique values
print(unique_values_mental)
```

```{r}
# Find elements that are common in both vectors
common_elements <- intersect(unique_values_mental, unique_values_happiness)

# Print the common elements
print(common_elements)
```

```{r}
Happiness_Data <- Happiness_Data[Happiness_Data$`Country name` %in% common_elements, ]

```

```{r}
Mental_Health_Data <- Mental_Health_Data[Mental_Health_Data$`Entity` %in% common_elements, ]
```

```{r}
unique_years_happy <- unique(Happiness_Data$year)
print(unique_years_happy)
```

```{r}
Mental_Health_Data <- Mental_Health_Data[Mental_Health_Data$Year %in% unique_years_happy, ]
```

```{r}
glimpse(Mental_Health_Data)
```

```{r}
glimpse(Happiness_Data)
```

```{r}
# Merge the datasets based on "Entity" and "Year"
merged_data <- merge(Mental_Health_Data, Happiness_Data, by.x = c("Entity", "Year"), by.y = c("Country name", "year"))


# Check the structure of the merged dataset
str(merged_data)
```

```{r}
glimpse(merged_data)
```

```{r}
# remove columns that are unreadbale and full of NA values
merged_data <- merged_data %>%
    select(-...11, -`Unique Countries`)
```

```{r}
glimpse(merged_data)
```

```{r}
# drop all rows that have NA values
# merged_data <- na.omit(merged_data)

glimpse(merged_data)
```

```{r}
#  Provide sample size, number of variables, and data source.

sample_size_merged_data <- nrow(merged_data)
num_variables_merged_data <- ncol(merged_data)

print(sample_size_merged_data)
print(num_variables_merged_data)
# ADD DATA SOURCE HERE
```

```{r}
#  Include a brief description of how the data was collected.

# i have no idea
```

-    Clearly state the research questions.

    -   To what extent do happiness metrics (predictor variables: log GDP per capita, social support, life expectancy, freedom, perception of corrpution, generosity) predict mental health disorders, specifically, depression (outcome variable: continuous) in countries?

    -   To what extent do mental health disorders and happiness metrics (predictor variables: schizophrenia, bipolar disorder, eating disorders, anxiety disorders, depression, log GDP per capita, social support, freedom, perception of corrpution, generosity) predict life ladder (outcome variable: categorical)?

```{r}
# Define a function to calculate descriptive statistics
calculate_descriptive_stats <- function(data) {
    # Get the names of numeric columns
    numeric_columns <- names(data)[sapply(data, is.numeric)]

    # Initialize a list to store results
    results <- list()

    # Loop through numeric columns
    for (col in numeric_columns) {
        stats <- summary(data[[col]])
        mean_val <- mean(data[[col]])
        median_val <- median(data[[col]])
        sd_val <- sd(data[[col]])
        ci_val <- t.test(data[[col]])$conf.int

        results[[col]] <- list(
            Summary = stats,
            Mean = mean_val,
            Median = median_val,
            StandardDeviation = sd_val,
            ConfidenceInterval = ci_val
        )
    }

    return(results)
}
```

```{r}
descriptive_stats <- calculate_descriptive_stats(merged_data)

for (col in names(descriptive_stats)) {
    cat("Descriptive Statistics for", col, ":\n")
    cat("- Summary:\n")
    print(descriptive_stats[[col]]$Summary)
    cat("- Mean:", descriptive_stats[[col]]$Mean, "\n")
    cat("- Median:", descriptive_stats[[col]]$Median, "\n")
    cat("- Standard Deviation:", descriptive_stats[[col]]$StandardDeviation, "\n")
    cat("- 95% Confidence Interval:", descriptive_stats[[col]]$ConfidenceInterval, "\n\n")
}
```

```{r}
# Create exploratory plots (tables or figures).
library(ggplot2)

ggplot(merged_data, aes(x = `Life Ladder`, y = `Log GDP per capita`)) +
    geom_point() +
    labs(x = "Life Ladder", y = "Log GDP per capita", title = "Scatter Plot")
```

```{r}
# Provide clear descriptions of findings.
# Like what are we doing here? I'm not even sure.
```

```{r}
detailed_variable_types <- sapply(merged_data, function(x) typeof(x))

# Print the detailed variable types
print(detailed_variable_types)
```

```{r}
# Entity -> categorical varaible for country?
# Life ladder -> categorical variable
# Year -> continous variable
# rest of varaibles are basically continous variables
# test push
# for vscode
View(merged_data)
```

```{r}
hist(merged_data$`Life Ladder`)
```

```{r}
# change life ladder to categorical variable 
median_life_ladder <- median(merged_data$`Life Ladder`)
# Create a new variable based on the median
merged_data$`Lifestyle Category` <- ifelse(merged_data$`Life Ladder` >= median_life_ladder, 1, 0)

# rename levels with labelss
merged_data$`Lifestyle Category` <- factor(merged_data$`Lifestyle Category`, levels = c(0, 1), labels = c("Below Average Lifestyle", "Above Average Lifestyle"))

```

```{r}

# Scatter plot for 'Depression (%)' vs 'Log GDP per capita'
ggplot(merged_data, aes(x = `Depression (%)`, y = `Log GDP per capita`)) +
  geom_point() +
  labs(title = 'Scatter Plot: Depression (%) vs Log GDP per capita',
       x = 'Depression (%)',
       y = 'Log GDP per capita') +
  theme_minimal()

# Scatter plot for 'Depression (%)' vs 'Social support'
ggplot(merged_data, aes(x = `Depression (%)`, y = `Social support`)) +
  geom_point() +
  labs(title = 'Scatter Plot: Depression (%) vs Social support',
       x = 'Depression (%)',
       y = 'Social support') +
  theme_minimal()

# Scatter plot for 'Depression (%)' vs 'Healthy life expectancy at birth'
ggplot(merged_data, aes(x = `Depression (%)`, y = `Healthy life expectancy at birth`)) +
  geom_point() +
  labs(title = 'Scatter Plot: Depression (%) vs Healthy life expectancy at birth',
       x = 'Depression (%)',
       y = 'Healthy life expectancy at birth') +
  theme_minimal()

# Scatter plot for 'Depression (%)' vs 'Freedom to make life choices'
ggplot(merged_data, aes(x = `Depression (%)`, y = `Freedom to make life choices`)) +
  geom_point() +
  labs(title = 'Scatter Plot: Depression (%) vs Freedom to make life choices',
       x = 'Depression (%)',
       y = 'Freedom to make life choices') +
  theme_minimal()

# Scatter plot for 'Depression (%)' vs 'Generosity'
ggplot(merged_data, aes(x = `Depression (%)`, y = Generosity)) +
  geom_point() +
  labs(title = 'Scatter Plot: Depression (%) vs Generosity',
       x = 'Depression (%)',
       y = 'Generosity') +
  theme_minimal()

# Scatter plot for 'Depression (%)' vs 'Perceptions of corruption'
ggplot(merged_data, aes(x = `Depression (%)`, y = `Perceptions of corruption`)) +
  geom_point() +
  labs(title = 'Scatter Plot: Depression (%) vs Perceptions of corruption',
       x = 'Depression (%)',
       y = 'Perceptions of corruption') +
  theme_minimal()

```
