---
echo: FALSE
editor: visual
message: FALSE
warning: FALSE
output:
 pdf_document:
  latex_engine: xelatex
 top-margin: 0.5in
geometry: margin = 2cm
format: pdf
---

## "Happiness and Mental Health Disorders \| Exploratory Data Analysis \| Aarya Desai \| Jeremy Tan \| Cassie Kang \| Osama Ahmed"

**Data Overview**: This assignment was a continuation of a research proposal submitted which outlined two potential datasets that we could use and further explore to ask two research questions. Based on the feedback, this assignment will be dedicated to conducting exploratory data analysis on the Happiness and Mental Health Dataset chosen from the previous report. This dataset is a combination of two datasets, namely, the World Happiness Dataset, 2022 retrieved from the official World Happiness Report website which conducts a yearly survey and analysis of world happiness metrics across the globe (i.e., each country), including different variables such as, life ladder, log GDP per capita, social support, life expectancy at birth, freedom, perception of corruption, generosity, positive affect and negative affect. The second dataset used was a World Mental Health Disorders dataset retrieved from the University of Oxford website containing datasets related to mental health metrics across the globe (i.e., each country) spanning over multiple years (). The latter dataset included metrics for mental health disorders recorded and analyzed on a Likert Scale (1 = Least Like to have given mental health disorder to 5 = Most likely to have given mental health disorder), and included this metric for the following disorders: Alcohol Use Disorders, Drug Use Disorders, Schizophrenia, Bipolar Disorders, Anxiety Disorders, Eating Disorders, and Depression. The resultant dataset had 1462 observations and 16 unique variables. These variables could be categorized as follows:

```{r, echo=FALSE, message=FALSE, results='hide'}
# Load the xtable package
library(xtable)

# Create a data frame for variable categorization
variable_categories <- data.frame(
  "Category" = c(
    "Subjective Well-being", "Economic Indicators", "Health and Life Expectancy",
    "Freedom and Governance", "Mental Health Disorders"
  ),
  "Variables" = c(
    "Life Ladder, Positive Affect, Negative Affect",
    "Log GDP per capita, Social Support",
    "Life Expectancy at Birth",
    "Freedom, Perception of Corruption, Generosity",
    "Alcohol Use Disorders, Drug Use Disorders, Schizophrenia, Bipolar Disorders, Anxiety Disorders, Eating Disorders, Depression"
  )
)

# Create the LaTeX table
variable_category_table <- xtable(variable_categories, caption = "Variable Categories", label = "tab:variable_categories")

# Print the table
print(variable_category_table)
```

```{r, echo=FALSE, message=FALSE, results='hide'}
cat("\\begin{table}[ht]\n")
cat("\\centering\n")
cat("\\begin{tabular}{rll}\n")
cat("  \\hline\n")
cat(" & Category & Variables \\\\ \n") 
cat("  \\hline\n")
cat("1 & Subjective Well-being & Life Ladder, Positive Affect, Negative Affect \\\\ \n")
cat("2 & Economic Indicators & Log GDP per capita, Social Support \\\\ \n")
cat("3 & Health and Life Expectancy & Life Expectancy at Birth \\\\ \n")
cat("4 & Freedom and Governance & Freedom, Perception of Corruption, Generosity \\\\ \n")
cat("5 & Mental Health Disorders & Alcohol Use Disorders, Drug Use Disorders, Schizophrenia, Bipolar Disorders, Anxiety Disorders, Eating Disorders, Depression \\\\ \n")
cat("   \\hline\n")
cat("\\end{tabular}\n")
cat("\\caption{Variable Categories} \n")
cat("\\label{tab:variable_categories}\n")
cat("\\end{table}\n")
```

The following is also a brief overview of the descriptive statistics for each of these variables, including the mean, median, standard deviation, and confidence intervals of each.

```{r, echo=FALSE, results='asis', message=FALSE}
cat("\\begin{table}[ht]\n")
cat("\\centering\n")
cat("\\begin{tabular}{rlllll}\n")
cat("  \\hline\n")
cat(" &  & Mean & Median & Standard Deviation & Confidence Interval \\\\ \n")
cat("  \\hline\n")
cat("1 & Year & 2011.813 & 2012 & 3.440 & 2011.636, 2011.989 \\\\ \n")
cat("  2 & Schizophrenia (\\%) & 0.212 & 0.200 & 0.045 & 0.210, 0.214 \\\\ \n")
cat("  3 & Bipolar disorder (\\%) & 0.743 & 0.708 & 0.151 & 0.735, 0.751 \\\\ \n")
cat("  4 & Eating disorders (\\%) & 0.263 & 0.201 & 0.172 & 0.254, 0.272 \\\\ \n")
cat("  5 & Anxiety disorders (\\%) & 4.008 & 3.556 & 1.277 & 3.942, 4.073 \\\\ \n")
cat("  6 & Drug use disorders (\\%) & 0.879 & 0.683 & 0.521 & 0.852, 0.906 \\\\ \n")
cat("  7 & Depression (\\%) & 3.460 & 3.479 & 0.641 & 3.427, 3.493 \\\\ \n")
cat("  8 & Alcohol use disorders (\\%) & 1.668 & 1.510 & 0.994 & 1.617, 1.719 \\\\ \n")
cat("  9 & Life Ladder & 5.448 & 5.347 & 1.142 & 5.389, 5.506 \\\\ \n")
cat("  10 & Log GDP per capita &  &  &  & 9.280, 9.399 \\\\ \n")
cat("  11 & Social support &  &  &  & 0.806, 0.818 \\\\ \n")
cat("  12 & Healthy life expectancy at birth & 62.750 & 64.620 & 7.202 & 62.381, 63.120 \\\\ \n")
cat("  13 & Freedom to make life choices &  &  &  & 0.726, 0.741 \\\\ \n")
cat("  14 & Generosity &  &  &  & -0.006, 0.012 \\\\ \n")
cat("  15 & Perceptions of corruption &  &  &  & 0.742, 0.762 \\\\ \n")
cat("  16 & Positive affect &  &  &  & 0.650, 0.661 \\\\ \n")
cat("  17 & Negative affect &  &  &  & 0.258, 0.267 \\\\ \n")
cat("   \\hline\n")
cat("\\end{tabular}\n")
cat("\\caption{Descriptive Statistics}\n")
cat("\\label{tab:descriptive_stats}\n")
cat("\\end{table}\n")
```

**Primary relationships of interest:** These datasets were then merged on corresponding Country and Year in each dataset. Based on this merged dataset, our team wanted to investigate the following research questions:

1.  To what extent do happiness metrics (predictor variables: log GDP per capita, social support, life expectancy, freedom, perception of corrpution, generosity) predict mental health disorders, specifically, depression (outcome variable: continuous) in countries?
2.  To what extent do mental health disorders and happiness metrics (predictor variables: schizophrenia, bipolar disorder, eating disorders, anxiety disorders, depression, log GDP per capita, social support, freedom, perception of corruption, generosity) predict life ladder (outcome variable: categorical)?

For the second research question, a binary outcome variable needed to be calculated from the original, continuous variable Life Ladder. This was to gain a more discerning outlook on the influence of various variables on how *good* or *bad* a countries' overall quality of life is. Therefore, for the same, Life Ladder was converted into a categorical variable named Lifestyle Category with two levels: Subpar Quality of Life and Satisfactory Quality of Life. This was done by assigning the prior factor to observations falling below the median of Life Ladder and the latter for those observations with a Life Ladder value above the median (*Mdn* = 5.35).

Figure 1. Life Ladder Distribution

```{r, echo=FALSE, results='hide', message=FALSE}
library(dplyr)
library(readxl)
```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
Happiness_Data <- read_xls("data/DataFortable2.1WHR2023.xls")
Mental_Health_Data <- read_xlsx("data/Mental health Depression disorder Data.xlsx")
```

```{r, echo=FALSE, results='hide', message=FALSE}
glimpse(Happiness_Data)
```

```{r, echo=FALSE, results='hide', message=FALSE}
glimpse(Mental_Health_Data)
```

```{r, echo=FALSE, results='hide', message=FALSE}
# Show unique values of a column
unique_values_happiness <- unique(Happiness_Data$`Country name`)

# Print the unique values
print(unique_values_happiness)
```

```{r, echo=FALSE, results='hide', message=FALSE}
# Show unique values of a column
unique_values_mental <- unique(Mental_Health_Data$Entity)

# Print the unique values
print(unique_values_mental)
```

```{r, echo=FALSE, results='hide', message=FALSE}
# Find elements that are common in both vectors
common_elements <- intersect(unique_values_mental, unique_values_happiness)

# Print the common elements
print(common_elements)
```

```{r, echo=FALSE, results='hide', message=FALSE}
Happiness_Data <- Happiness_Data[Happiness_Data$`Country name` %in% common_elements, ]

```

```{r, echo=FALSE, results='hide', message=FALSE}
Mental_Health_Data <- Mental_Health_Data[Mental_Health_Data$`Entity` %in% common_elements, ]
```

```{r, echo=FALSE, results='hide', message=FALSE}
unique_years_happy <- unique(Happiness_Data$year)
print(unique_years_happy)
```

```{r, echo=FALSE, results='hide', message=FALSE}
Mental_Health_Data <- Mental_Health_Data[Mental_Health_Data$Year %in% unique_years_happy, ]
```

```{r, echo=FALSE, results='hide', message=FALSE}
glimpse(Mental_Health_Data)
```

```{r, echo=FALSE, results='hide', message=FALSE}
glimpse(Happiness_Data)
```

```{r, echo=FALSE, results='hide', message=FALSE}
# Merge the datasets based on "Entity" and "Year"
merged_data <- merge(Mental_Health_Data, Happiness_Data, by.x = c("Entity", "Year"), by.y = c("Country name", "year"))


# Check the structure of the merged dataset
str(merged_data)
```

```{r, echo=FALSE, results='hide', message=FALSE}
glimpse(merged_data)
```

```{r, echo=FALSE, results='hide', message=FALSE}
# remove columns that are unreadbale and full of NA values
merged_data <- merged_data %>%
    select(-...11, -`Unique Countries`)
```

```{r, echo=FALSE, results='hide', message=FALSE}
glimpse(merged_data)
```

```{r, echo=FALSE, results='hide', message=FALSE}
# drop all rows that have NA values
# merged_data <- na.omit(merged_data)

glimpse(merged_data)
```

```{r, echo=FALSE, results='hide', message=FALSE}
#  Provide sample size, number of variables, and data source.

sample_size_merged_data <- nrow(merged_data)
num_variables_merged_data <- ncol(merged_data)

print(sample_size_merged_data)
print(num_variables_merged_data)
# ADD DATA SOURCE HERE
```

```{r, echo=FALSE, results='hide', message=FALSE}
#  Include a brief description of how the data was collected.

# i have no idea
```

```{r, echo=FALSE, results='hide', message=FALSE}
# Define a function to calculate descriptive statistics
calculate_descriptive_stats <- function(data) {
    # Get the names of numeric columns
    numeric_columns <- names(data)[sapply(data, is.numeric)]

    # Initialize a list to store results
    results <- list()

    # Loop through numeric columns
    for (col in numeric_columns) {
        stats <- summary(data[[col]])
        mean_val <- mean(data[[col]])
        median_val <- median(data[[col]])
        sd_val <- sd(data[[col]])
        ci_val <- t.test(data[[col]])$conf.int

        results[[col]] <- list(
            Summary = stats,
            Mean = mean_val,
            Median = median_val,
            StandardDeviation = sd_val,
            ConfidenceInterval = ci_val
        )
    }

    return(results)
}
```

```{r, echo=FALSE,results='hide', message=FALSE}
# Load the xtable package
library(xtable)

# Calculate descriptive statistics
descriptive_stats <- calculate_descriptive_stats(merged_data)

# Create a data frame to hold the summary statistics
summary_df <- data.frame(
  Variable = character(0),
  Mean = numeric(0),
  Median = numeric(0),
  StandardDeviation = numeric(0),
  ConfidenceInterval = character(0)
)

# Populate the summary data frame
for (col in names(descriptive_stats)) {
  summary_df <- rbind(summary_df, c(col, 
                                    descriptive_stats[[col]]$Mean,
                                    descriptive_stats[[col]]$Median,
                                    descriptive_stats[[col]]$StandardDeviation,
                                    ifelse(length(descriptive_stats[[col]]$ConfidenceInterval) > 1,
                                           paste(descriptive_stats[[col]]$ConfidenceInterval, collapse = ", "),
                                           descriptive_stats[[col]]$ConfidenceInterval)
  ))
}

# Create the LaTeX table
table <- xtable(summary_df, caption = "Descriptive Statistics", label = "tab:descriptive_stats")
```

```{r, echo=FALSE, results='hide', message=FALSE, fig.show='hide'}
# Create exploratory plots (tables or figures).
library(ggplot2)

ggplot(merged_data, aes(x = `Life Ladder`, y = `Log GDP per capita`)) +
    geom_point() +
    labs(x = "Life Ladder", y = "Log GDP per capita", title = "Scatter Plot")
```

```{r, echo=FALSE, results='hide', message=FALSE}
# Provide clear descriptions of findings.
# Like what are we doing here? I'm not even sure.
```

```{r, echo=FALSE, results='hide', message=FALSE}
detailed_variable_types <- sapply(merged_data, function(x) typeof(x))

# Print the detailed variable types
print(detailed_variable_types)
```

```{r, echo=FALSE, results='hide', message=FALSE}
# Entity -> categorical varaible for country?
# Life ladder -> categorical variable
# Year -> continous variable
# rest of varaibles are basically continous variables
# test push
# for vscode
# View(merged_data)
```

```{r, echo=FALSE,  results='hide',fig.height=2, fig.width=4, message=FALSE, warning=FALSE}
# Create a histogram with density plot overlay
ggplot(merged_data, aes(x = `Life Ladder`)) +
  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black", aes(y = ..density..)) +
  geom_density(color = "navy", size = 1) +
  labs(x = "Life Ladder",
       y = "Density") +
  theme_minimal()

```

```{r, echo=FALSE, fig.height=4, fig.width=8, message=FALSE, warning=FALSE}
# change life ladder to categorical variable 
median_life_ladder <- median(merged_data$`Life Ladder`)
# Create a new variable based on the median
merged_data$`Lifestyle Category` <- ifelse(merged_data$`Life Ladder` >= median_life_ladder, 1, 0)

# rename levels with labelss
merged_data$`Lifestyle Category` <- factor(merged_data$`Lifestyle Category`, levels = c(0, 1), labels = c("Subpar QoL", "Satisfactory QoL"))

```

**Outcome variables**

**Research Question 1 (RQ1):** The following are plots to visually represent the predictors variables for this research question against depression.

Figure 2. RQ1 predictor variables against the outcome variable depression

```{r, echo=FALSE, results='hide', fig.height=4, fig.width=6, message=FALSE}

# Load the gridExtra package
library(gridExtra)
library(ggplot2)
library(cowplot)


# Create scatter plots
plots <- list(
  ggplot(merged_data, aes(x = `Depression (%)`, y = `Log GDP per capita`)) +
    geom_point(color = "blue", alpha = 0.6) +
    geom_smooth(method = "lm", color = "lightblue") +
    labs(title = 'Scatter Plot: Depression (%) vs Log GDP per capita',
         x = 'Depression (%)',
         y = 'Log GDP per capita') +
    theme_minimal()+
    theme(
        plot.title = element_text(size = rel(0.5)),
        axis.title.x = element_text(size = rel(0.6)),
        axis.title.y = element_text(size = rel(0.6)),
        axis.text.x = element_text(size = rel(0.6)),
        axis.text.y = element_text(size = rel(0.6))
    ),

  ggplot(merged_data, aes(x = `Depression (%)`, y = `Social support`)) +
    geom_point(color = "green", alpha = 0.6) +
    geom_smooth(method = "lm", color = "white") +
    labs(title = 'Scatter Plot: Depression (%) vs Social support',
         x = 'Depression (%)',
         y = 'Social support') +
    theme_minimal()+
    theme(
        plot.title = element_text(size = rel(0.5)),
        axis.title.x = element_text(size = rel(0.6)),
        axis.title.y = element_text(size = rel(0.6)),
        axis.text.x = element_text(size = rel(0.6)),
        axis.text.y = element_text(size = rel(0.6))
    ),

  ggplot(merged_data, aes(x = `Depression (%)`, y = `Healthy life expectancy at birth`)) +
    geom_point(color = "purple", alpha = 0.6) +
    geom_smooth(method = "lm", color = "violet") +
    labs(title = 'Scatter Plot: Depression (%) vs Healthy life expectancy at birth',
         x = 'Depression (%)',
         y = 'Healthy life expectancy at birth') +
    theme_minimal()+
    theme(
        plot.title = element_text(size = rel(0.5)),
        axis.title.x = element_text(size = rel(0.6)),
        axis.title.y = element_text(size = rel(0.6)),
        axis.text.x = element_text(size = rel(0.6)),
        axis.text.y = element_text(size = rel(0.6))
    ),

  ggplot(merged_data, aes(x = `Depression (%)`, y = `Freedom to make life choices`)) +
    geom_point(color = "orange", alpha = 0.6) +
    geom_smooth(method = "lm", color = "yellow") +
    labs(title = 'Scatter Plot: Depression (%) vs Freedom to make life choices',
         x = 'Depression (%)',
         y = 'Freedom to make life choices') +
    theme_minimal()+
    theme(
        plot.title = element_text(size = rel(0.5)),
        axis.title.x = element_text(size = rel(0.6)),
        axis.title.y = element_text(size = rel(0.6)),
        axis.text.x = element_text(size = rel(0.6)),
        axis.text.y = element_text(size = rel(0.6))
    ),

  ggplot(merged_data, aes(x = `Depression (%)`, y = Generosity)) +
    geom_point(color = "brown", alpha = 0.6) +
    geom_smooth(method = "lm", color = "black") +
    labs(title = 'Scatter Plot: Depression (%) vs Generosity',
         x = 'Depression (%)',
         y = 'Generosity') +
    theme_minimal()+
    theme(
        plot.title = element_text(size = rel(0.5)),
        axis.title.x = element_text(size = rel(0.6)),
        axis.title.y = element_text(size = rel(0.6)),
        axis.text.x = element_text(size = rel(0.6)),
        axis.text.y = element_text(size = rel(0.6))
    ),

  ggplot(merged_data, aes(x = `Depression (%)`, y = `Perceptions of corruption`)) +
    geom_point(color = "pink", alpha = 0.6) +
    geom_smooth(method = "lm", color = "red") +
    labs(title = 'Scatter Plot: Depression (%) vs Perceptions of corruption',
         x = 'Depression (%)',
         y = 'Perceptions of corruption') +
    theme_minimal()+
    theme(
        plot.title = element_text(size = rel(0.5)),
        axis.title.x = element_text(size = rel(0.6)),
        axis.title.y = element_text(size = rel(0.6)),
        axis.text.x = element_text(size = rel(0.6)),
        axis.text.y = element_text(size = rel(0.6))
    )
)

# Arrange the plots in a grid
arranged_plots <- grid.arrange(grobs = plots, ncol = 2)
arranged_plots

```

As seen in the figures above, the data is showing distinct patterns in a clustered fashion. This is because the data is being inherently group based on the country, changing very minimally by the year. Therefore, to have a better understanding of the relationship without having sub-trends in the graphs, the data was transformed, where the mean of each variable was calculated per country in the dataset. Though this reduced the number of observations from 1462 to 152, this was a necessary step required to gain some more understandable exploratory information regarding the data without accounting for the confound of the Year variable. Below are the new plots with the averaged variables plotted against depression.

Figure 3. RQ1 predictor variables against the outcome variable depression, with variables grouped and averaged by country

```{r, echo=FALSE, results='hide', fig.height=4, fig.width=6, message=FALSE}
# Group the data by the 'Entity' (countries) column
grouped_data <- merged_data %>%
  group_by(Entity)

# Calculate the average for the selected variables
averaged_data <- grouped_data %>%
  summarise(
    `Perceptions of corruption` = mean(`Perceptions of corruption`, na.rm = TRUE),
    Generosity = mean(Generosity , na.rm = TRUE),
    `Freedom to make life choices` = mean(`Freedom to make life choices`, na.rm = TRUE),
    `Healthy life expectancy at birth` = mean(`Healthy life expectancy at birth`, na.rm = TRUE),
    `Social support` = mean(`Social support`, na.rm = TRUE),
    `Log GDP per capita` = mean(`Log GDP per capita`, na.rm = TRUE),
    `Depression (%)` = mean(`Depression (%)`, na.rm = TRUE)
  )

# View the resulting dataset with averages
# View(averaged_data)
```

```{r, echo=FALSE, results='hide', fig.height=4, fig.width=6, message=FALSE}

# Load the required libraries
library(gridExtra)
library(ggplot2)

# Create scatter plots
plots_new <- list(
  ggplot(averaged_data, aes(x = `Depression (%)`, y = `Log GDP per capita`)) +
    geom_point(color = "blue", alpha = 0.6) +
    geom_smooth(method = "lm", color = "lightblue") +
    labs(title = 'Scatter Plot: Depression (%) vs Log GDP per capita',
         x = 'Depression (%)',
         y = 'Log GDP per capita') +
    theme_minimal()+
    theme(
        plot.title = element_text(size = rel(0.5)),
        axis.title.x = element_text(size = rel(0.6)),
        axis.title.y = element_text(size = rel(0.6)),
        axis.text.x = element_text(size = rel(0.6)),
        axis.text.y = element_text(size = rel(0.6))
    ),

  ggplot(averaged_data, aes(x = `Depression (%)`, y = `Social support`)) +
    geom_point(color = "green", alpha = 0.6) +
    geom_smooth(method = "lm", color = "white") +
    labs(title = 'Scatter Plot: Depression (%) vs Social support',
         x = 'Depression (%)',
         y = 'Social support') +
    theme_minimal()+
    theme(
        plot.title = element_text(size = rel(0.5)),
        axis.title.x = element_text(size = rel(0.6)),
        axis.title.y = element_text(size = rel(0.6)),
        axis.text.x = element_text(size = rel(0.6)),
        axis.text.y = element_text(size = rel(0.6))
    ),

  ggplot(averaged_data, aes(x = `Depression (%)`, y = `Healthy life expectancy at birth`)) +
    geom_point(color = "purple", alpha = 0.6) +
    geom_smooth(method = "lm", color = "violet") +
    labs(title = 'Scatter Plot: Depression (%) vs Healthy life expectancy at birth',
         x = 'Depression (%)',
         y = 'Healthy life expectancy at birth') +
    theme_minimal()+
    theme(
        plot.title = element_text(size = rel(0.5)),
        axis.title.x = element_text(size = rel(0.6)),
        axis.title.y = element_text(size = rel(0.6)),
        axis.text.x = element_text(size = rel(0.6)),
        axis.text.y = element_text(size = rel(0.6))
    ),

  ggplot(averaged_data, aes(x = `Depression (%)`, y = `Freedom to make life choices`)) +
    geom_point(color = "orange", alpha = 0.6) +
    geom_smooth(method = "lm", color = "yellow") +
    labs(title = 'Scatter Plot: Depression (%) vs Freedom to make life choices',
         x = 'Depression (%)',
         y = 'Freedom to make life choices') +
    theme_minimal()+
    theme(
        plot.title = element_text(size = rel(0.5)),
        axis.title.x = element_text(size = rel(0.6)),
        axis.title.y = element_text(size = rel(0.6)),
        axis.text.x = element_text(size = rel(0.6)),
        axis.text.y = element_text(size = rel(0.6))
    ),

  ggplot(averaged_data, aes(x = `Depression (%)`, y = Generosity)) +
    geom_point(color = "brown", alpha = 0.6) +
    geom_smooth(method = "lm", color = "black") +
    labs(title = 'Scatter Plot: Depression (%) vs Generosity',
         x = 'Depression (%)',
         y = 'Generosity') +
    theme_minimal()+
    theme(
        plot.title = element_text(size = rel(0.5)),
        axis.title.x = element_text(size = rel(0.6)),
        axis.title.y = element_text(size = rel(0.6)),
        axis.text.x = element_text(size = rel(0.6)),
        axis.text.y = element_text(size = rel(0.6))
    ),

  ggplot(averaged_data, aes(x = `Depression (%)`, y = `Perceptions of corruption`)) +
    geom_point(color = "pink", alpha = 0.6) +
    geom_smooth(method = "lm", color = "red") +
    labs(title = 'Scatter Plot: Depression (%) vs Perceptions of corruption',
         x = 'Depression (%)',
         y = 'Perceptions of corruption') +
    theme_minimal()+
    theme(
        plot.title = element_text(size = rel(0.5)),
        axis.title.x = element_text(size = rel(0.6)),
        axis.title.y = element_text(size = rel(0.6)),
        axis.text.x = element_text(size = rel(0.6)),
        axis.text.y = element_text(size = rel(0.6))
    )
)

# Arrange the plots in a grid
arranged_plots <- grid.arrange(grobs = plots_new, ncol = 2)

arranged_plots


```

These grid plots reveal that the connections between depression and the chosen predictor variables are not notably pronounced. The only variable that shows a weak, negative relationship with depression is 'Perception of Corruption'. On the other hand, the remaining variables exhibit relatively flat relationship lines, indicating a more subtle correlation, and providing us with limited additional insights into their interrelationship with depression. Therefore, these plots do not provide much insightful information regarding RQ2.

Lastly, for some exploratory understanding of the general distribution of depression rates across the globe, a map plot was made to see the variation in depression rates between countries (Figure 4). As seen in the figure, there is a notable variation in the depression metric between countries. Therefore, our model may benefit from using Country as a factor or confound, to gain a more in-depth understanding of RQ1.

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE, fig.show='hide'}
#Factored bar plot for 'Depression (%)' by 'Entity'
ggplot(averaged_data, aes(x = reorder(Entity, `Depression (%)`), y = `Depression (%)`)) +
    geom_bar(stat = "identity") +
    coord_flip() +  # Flip coordinates to display a horizontal bar chart
    labs(
        title = "Average Depression Rate by Country",
        x = "Country",
        y = "Average Depression Rate (%)"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
          axis.text.y = element_text(size = 2))

# Factored bar plot for 'Depression (%)' by 'Lifestyle Category'
ggplot(merged_data, aes(x = `Lifestyle Category`, y = `Depression (%)`, fill = `Lifestyle Category`)) +
  geom_bar(stat = "summary", fun = "mean") +
  labs(title = 'Factored Bar Plot: Depression (%) by Lifestyle Category',
       x = 'Lifestyle Category',
       y = 'Mean Depression Rate (%)') +
  theme_minimal()
```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
#install.packages("mapdata")
```

Figure 4. Heatmap to visualize depression rates across countries.

```{r, echo=FALSE, results='hide', fig.height=4, fig.width=6, message=FALSE}
# Load the required libraries
library(ggplot2)
library(maps)
library(mapdata)

# Assuming `averaged_data` contains the country names and depression rates

# Get world map data
world_map <- map_data("world")

# Perform a left join to keep only the countries in averaged_data
merged_map_data <- merge(world_map, averaged_data, by.x = "region", by.y = "Entity", all.x = TRUE)

ggplot(merged_map_data, aes(x = long, y = lat, group = group, fill = `Depression (%)`)) +
  geom_polygon() +
  borders("world", xlim = c(-180, 180), ylim = c(-90, 90), colour = "white", fill = NA) +  # Add world borders
  scale_fill_gradient(low = "blue", high = "red", name = "Depression (%)") +
  labs(title = "Depression Rate Choropleth Map",
       x = "Longitude",
       y = "Latitude") +
  theme(legend.position = "bottom") +
  coord_fixed(1.3)
```

**Research Question 2 (RQ2):** To gain a better understanding of the relationship between the binary outcome variable, Lifestyle Category, and the continuous predictor variables chosen for RQ2, bar plots to show the difference in Subpar vs Satisfactory Quality of Life (QoL) were plotted (Figure 5).

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE, fig.show='hide'}
# Create a new factor variable for Depression rate categories
min_val <- min(merged_data$`Depression (%)`, na.rm = TRUE) # excluding NA values
max_val <- max(merged_data$`Depression (%)`, na.rm = TRUE)

# Create a new factor variable for Depression rate categories
merged_data$`Depression Category` <- cut(merged_data$`Depression (%)`,
                                         breaks = c(min_val, 2, 4, max_val),
                                         labels = c("relatively Low", "relatively Moderate", "relatively High"),
                                         right = FALSE) 

# Factored bar plot for 'Life Ladder' by 'Depression Category'
ggplot(merged_data, aes(x = `Depression Category`, y = `Life Ladder`, fill = `Depression Category`)) +
  geom_bar(stat = "summary", fun = "mean") +
  labs(title = 'Factored Bar Plot: Life Ladder by Depression Category',
       x = 'Depression Percentage',
       y = 'Mean Life Ladder Score') +
  theme_minimal()
```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE, fig.show='hide'}
# Scatter plot for 'Life Ladder' vs 'Continuous Predictor 1'
ggplot(merged_data, aes(x = `Depression (%)`, y = `Lifestyle Category`, color = `Lifestyle Category`)) +
  geom_boxplot() +
  labs(title = 'Box Plot: Depression by Life Ladder',
       x = 'Depression',
       y = 'Life Ladder') +
  theme_minimal()
```

Figure 5.

```{r, echo=FALSE, results='hide', fig.height=7, fig.width=11, message=FALSE}
predictor_vars <- c("Schizophrenia (%)", "Bipolar disorder (%)", "Eating disorders (%)", 
                    "Anxiety disorders (%)", "Depression (%)", "Log GDP per capita", 
                    "Social support", "Freedom to make life choices", "Perceptions of corruption", "Generosity")

# Create a list to store the plots
plots_rq2 <- list()

# Loop through the predictor variables
for (var in predictor_vars) {
  if (is.numeric(merged_data[[var]])) {
    # For continuous predictors, create a box plot
    p <- ggplot(merged_data, aes(x = `Lifestyle Category`, y = .data[[var]], fill = `Lifestyle Category`)) +
      geom_boxplot() +
      labs(title = paste("Box Plot of", var, "by Lifestyle Category"),
           x = "Lifestyle Category",
           y = var) +
      theme_minimal() +
      theme(
        plot.title = element_text(size = rel(0.6)),
        axis.title.x = element_text(size = rel(0.5)),
        axis.title.y = element_text(size = rel(0.4)),
        axis.text.x = element_text(size = rel(0.5)),
        axis.text.y = element_text(size = rel(0.5)),
        legend.title = element_text(size = rel(0.3)),
        legend.text = element_text(size = rel(0.3))
      )
  } else {
    # For binary predictors, create a bar plot
    p <- ggplot(merged_data, aes(x = .data[[var]], fill = `Lifestyle Category`)) +
      geom_bar(position = "dodge") +
      labs(title = paste("Bar Plot of", var, "by Lifestyle Category"),
           x = var,
           y = "Frequency") +
      theme_minimal() +
      theme(
        plot.title = element_text(size = rel(0.6)),
        axis.title.x = element_text(size = rel(0.5)),
        axis.title.y = element_text(size = rel(0.4)),
        axis.text.x = element_text(size = rel(0.5)),
        axis.text.y = element_text(size = rel(0.5)),
        legend.title = element_text(size = rel(0.3)),
        legend.text = element_text(size = rel(0.3))
      )
  }
  plots_rq2[[var]] <- p
}

grid.arrange(grobs = plots_rq2, ncol = 2)


```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE, fig.show='hide'}
# Factored bar plot for 'Life Ladder' categorized by 'Lifestyle Category'
ggplot(merged_data, aes(x = `Lifestyle Category`, fill = `Lifestyle Category`)) +
    geom_bar() +
    labs(
        title = "Distribution of Life Ladder Categories",
        x = "Lifestyle Category",
        y = "Count"
    ) +
    theme_minimal()

# Factored bar plot for 'Lifestyle Category' vs 'Log GDP per capita'
ggplot(merged_data,aes(x = `Lifestyle Category`, y = `Log GDP per capita`, fill = `Lifestyle Category`)) +
  geom_bar(stat = "summary", fun = "mean") +
  labs(title = 'Lifestyle Category vs Log GDP per capita',
       x = 'Lifestyle Category',
       y = 'Mean Log GDP per capita') +
  theme_minimal()
```

**Final Variable Selection:**

###RQ1
In the modeling process for the first research question, our priori variable selection determined the following variables based on their established impact on happiness and well-being. These include Log GDP per capita, which is often associated with better living conditions and higher well-being. Social support, another key predictor, signifies the importance of strong social connections in mental well-being. Healthy life expectancy at birth, reflecting the availability of quality healthcare, and freedom to make life choices, indicating the role of personal freedoms in happiness, were also chosen. Generosity, denoting acts of kindness and generosity, and perceptions of corruption, as a measure of societal transparency, complete the variable selection.

###RQ2
For the modeling process in the second research question, our priori variable selection includes Schizophrenia, known for its significant impact on overall well-being, and Eating disorders, which often have a direct effect on mental health. Bipolar disorder, characterized by mood swings. Anxiety disorder, associated with happiness and life quality. We also selected Social support, recognizing its role as a fundamental factor in both happiness and mental well-being, and Log GDP per capita, an indicator of economic stability, were chosen as predictive variables. Additionally, Freedom, representing the freedom to make life choices, was included as an influential variable. These selections are grounded in their potential to influence both happiness and mental health outcomes, thus affecting our Life Ladder outcome variable.

**Other Characteristics**: In the analysis of the World Happiness Report and World Mental Health Data, it was essential to consider variables that would not be included in our model. These variables are "Country"/"Entity" and "Year."

Year is a categorical variable representing the years for which the data was collected. It has 13 unique values with varying repetitions. The "Country" variable represents the countries for which the data was collected. It is a categorical variable with 153 unique countries, each with varying repetitions.

**Potential Challenges**:

1.  Assumptions about Data Homogeneity: It is assumed that the data is uniform across countries and years, disregarding potential influences from these variables. This is crucial because factors like GDP and cultural nuances can significantly impact happiness metrics.

2.  Addressing NAs: There's a need to understand why there are missing values (NAs). This could be due to merging processes or a specific pattern (e.g., related to certain years, countries, or occurring randomly). It's important to address this issue, as it could affect the accuracy of the results or model.

3.  Binary Variable Creation: Creating a binary variable based on the median may not capture nuanced differences between groups. This might limit the depth of insights derived from the analysis.

4.  Limited Insights from EDA for RQ1: Initial exploratory data analysis (EDA) for Research Question 1 suggests weak relationships. This could potentially lead to a model with limited predictive power and insights.

5.  Promising Aspects of RQ2: Research Question 2 holds more promise, but there are still important considerations, starting with an examination of how the binary variable was derived.

6.  Multicollinearity Concerns: Given the nature of predictor variables in the World Happiness Dataset, there's a possibility of high correlation among them. This can lead to multicollinearity in regression models, complicating the interpretation of individual predictor effects on the outcome.
